# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

First we load the data
```{r}
data <- read.csv("activity.csv")

```


## What is mean total number of steps taken per day?

This is the histogram of the total number of steps per day
```{r fig.width=6, fig.height=6}
sumxdate<-unlist(by(data$steps,data$date,sum,na.rm=TRUE))
hist(sumxdate, xlab = "Total steps", ylab = "# of days", main = "")

```

The  mean total number of steps taken per day
```{r}
mean(sumxdate)
```
The  median total number of steps taken per day
```{r}
median(sumxdate)

```


## What is the average daily activity pattern?

```{r}
meanxint<-unlist(by(data$steps,data$interval,mean,na.rm=TRUE))
plot(meanxint, type='l',xlab = "Minutes", ylab = "Steps", main = "Average time series")

```

The interval corresponding to the maximum average number of steps is:
```{r}
which.max(meanxint)
```
First numbercorresponds to the hour (ex. 835 means 8:35)  
Second number corresponds to the actual interval index

## Imputing missing values

The total number of rows with missing step data is:
```{r}
sum(is.na(data[,1]))
```
We will fill missing values with the average number of steps in that interval

```{r}
dataf<-data
r<-rep(meanxint,length(split(data,data$date)))
dataf[is.na(dataf[,1]),1]<-r[is.na(dataf[,1])]
```
Now the total number rows with missing step data is:
```{r}
sum(is.na(dataf[,1]))
```

We plot the histogram of total steps per day and calculate mean and median as above with the new dataset:

```{r fig.width=6, fig.height=6}
sumxdate2<-unlist(by(dataf[,1],dataf[,2],sum,na.rm=TRUE))
hist(sumxdate2, xlab = "Total steps", ylab = "# of days", main = "")

mean(sumxdate2)

median(sumxdate2)

```

We can see that imputing missing data increased the total number of steps, as expected.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
Sys.setlocale("LC_TIME", "English")
dataf$week <- weekdays(as.POSIXlt(dataf$date))
dataf$day[dataf$week %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")]<-c("weekday")
dataf$day[dataf$week %in% c("Saturday","Sunday")]<-c("weekend")
dataf$day<-factor(dataf$day)
```

Plot of time series divided weekends and weekdays:
```{r}
library(lattice)
ag<-aggregate(steps ~ day + interval, data = dataf, FUN= "mean" )
xyplot(steps ~ interval | day, data = ag, layout = c(1, 2), type = 'l')

```

